### 2011-04-24 doc/Manual/BaconManual.tex
Started on paper.

\documentclass{article}
\begin{document}

\section{Introduction}

Bacon is a front end for OpenCL that provides extra convienience compared
to the raw OpenCL API and OpenCL C programming language.

Bacon is a data parallel extension to OpenCL C that adds the following
features:

 - Native 2d and 3d arrays with multi-dimensional addressing syntax.

 - Images as multi-dimensional arrays.

 - Support for "printf debugging".

 - Automatic C++ wrapper generation.

\section{Multi-Dimentional Arrays}

Support is provided for native 2D and 3D arrays. Some examples:

Declare a 2D array with eight rows and four columns:
$$\hbox{\tt array2d<int> table[8, 4];}$$

Assign the value 8 to the third row, second column of the array.
$$\hbox{\tt table[2, 1] = table.rows;}$$

\end{document}